On The Impact of Atoms of Confusion in JavaScript Code

### 2021-01-20 4_Result.tex
revisao da secao de study settings e resultados

\section{Results and Discussion}
\label{sec:results}

In this section we present the main findings of our research. We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}). After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.

\subsection{Results of the first study: a survey}
\label{sec:survey-resuts}

Our first study aims to investigate the impact of atoms of confusion while developers try to understand JavaScript code. We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers) and \emph{cognitive effort} (time necessary to provide an answer). As discussed in section~\ref{sec:survey-settings}, 140 participants answered the survey in full. Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects, according to their education level and years of programming experience, respectively.

\begin{figure}[htb]
      \centering
      \includegraphics[scale=.25]{images/degree.png}
      \caption{Participants' Education Level}\label{fig:degree}
  \end{figure}

  \begin{figure}[htb]
      \centering
      \includegraphics[scale=.25]{images/experience.png}
      \caption{Participants' Experience} \label{fig:xp}
  \end{figure}

It is possible to realize in Figure \ref{fig:xp} that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming. No subject reported to have never attended university. Out audience comprises experience developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}). Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%). Other 37\% of the respondents have between one and four years of experience. In this way, we can characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.

Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained simplified versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. We collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). Our final dataset consists of 70 complete Latin Squares, which means that our assessment consider a total of 140 answers to each pair of code snippets.

Regarding \emph{correctness}, Table~\ref{results_correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey. Accordingly, the version of seven code snippets without atoms of confusion present at least a 15\% improvement in answer correctness. In particular, \emph{Comma Operator} atom presents the highest impact on misunderstanding. 
Curiously, frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion. As the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when the atoms were not present. Also, the sample of answers where there was no atoms had almost no dispersion, which is a sign that the non-confusing code is easier to evaluate correctly.

@@

\begin{table}[htbp]
\caption{Average number of correct answers.}
\label{tab:difference-correctness}
\begin{tabular}{lccc} \toprule
 & \multicolumn{2}{c}{Average Number of Correct Answers} & \\ [0.1cm]
 Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
 Comma Operator          & 40 & 93  & +132 \\
 Post Increment          & 69 & 91  & + 31  \\
 Omitted Curly Braces    & 67 & 83  & +23 \\
 Assignment as Value     & 80 & 97  & +21 \\
 Implicit Predicate      & 83 & 97  & +16 \\
 Logic as Control Flow   & 59 & 68  & +15 \\
 Ternary Operator        & 86 & 94  & +9  \\
 Pre-Increment           & 71 & 76  & +7  \\
 Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
\end{tabular}
\end{table}

\begin{figure}[htb!]
\noindent
 \centering
 \includegraphics[scale=.20]{images/wrong.png}
 \caption{Number of wrong answers of each subject}
 \label{fig:boxplotcorrectness}
 \end{figure}

\rb{verificar a consistencia nas questoes de pesquisa mais especificas}
Regarding the first question we 
address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code. We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).

\rb{acho que valeria a pena colocar o teste de hipotese aqui}
\adriano{como fazemos/descrevemos o teste?}

Table \ref{tab:difference-time-taken} displays the results of the measurements of the average time taken to answer with and without atoms of confusion. Once again, the \emph{Comma Operator} was the construct whose removal had the greatest impact. The more expressive results in this measurement are the ones regarding \emph{Logic as Control Flow}, \emph{Implicit Predicate} and \emph{Omitted Curly Braces}, {\color{blue}as they are frequent in JavaScript libraries}.

\rb{n\~{a}o sei se devemos antecipar o resultado do estudo de minera\c c\~{a}o 
nessa se\c c\~{a}o. talvez nao comprometa tanto.}

Figure \ref{fig:timetoanswer} shows a box plot of the difference in time taken the predict the output of average time to assess the code snippets. Although the dispersion is smaller for non-confusing code blocks, the median time did not vary much.

\begin{table}[htbp]
\caption{Average number of correct answers.}
\label{tab:difference-time-taken}
\begin{tabular}{lccc} \toprule
 & \multicolumn{2}{c}{Average Time in Seconds to Answer} & \\ [0.1cm]
 Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
Comma Operator        & 60 & 21  & -65  \\
Logic as Control Flow & 85 & 49  & -42  \\
Implicit Predicate    & 33 & 24  & -27  \\
Omitted Curly Braces  & 43 & 31  & -27  \\
Assignment as Value   & 53 & 49  & -7   \\
Ternary Operator      & 42 & 42  & 0    \\
Post Increment        & 27 & 27  & 0    \\
Arithmetic as Logic   & 29 & 36  & +24  \\
Pre-Increment         & 34 & 49  & +44  \\
         \bottomrule
    \end{tabular}
\end{table}

\begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.20]{images/timeeach.png}
      \caption{Box plot of time to answer each snippet}
      \label{fig:timetoanswer}
  \end{figure}

Recalling our second research question of this survey (\emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}), the variance in the time to correctly predict the output of a snippet is significantly smaller. Moreover, since the number of answers is the same for each inspected atom, data from table \ref{results_time} shows that the time taken to answer questions without atoms of confusion is 20\% smaller. Not all atoms, though, took less time to predict the answer. In fact, for the \emph{Arithmetic as Logic} and the \emph{Pre-Increment} atoms, there was an increase of 24\% and 44\% in the time taken to answer questions wherein the atoms had been removed. For other two atoms, there is no difference whatsoever in the time taken. We only find in half of our atoms an increase in the time taken to answer. Therefore, even though we are not able to confirm hypotheses H.2 defined in Section 1, further inquiry would be necessary to discover the reason for some atoms taking even longer.

\rb{talvez tenhamos que descrever e testar as hipoteses. a descricao pode ser feita na secao de study settings, e o resultado dos testes pode aparecer aqui.}

% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

\subsection{Results of the second study: Interview}
\label{sec:interview-results}

In this section we present the results of two rounds of two rounds of interviews with practitioners who agreed to participate in the interviews.In the first interview, a total of 15 practitioners were surveyed. In this phase, we collected information regarding programming experience, familiarity with JavaScript, and we also collected their opinions regarding nine candidates to atoms of confusion. This was done by presenting them pairs of code snippets whose behavior was identical, and ask them to classify each snippets readability.

In the second round of interviews, ten of the fifteen developers who participated in the first round were shown another 8 candidates to atom of confusion whose frequency we observed during the repository mining phase. The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.

Supporting the results of the survey,Tables \ref{tab:interview-results1} and \ref{tab:interview-results2} show that for twelve out of the seventeen scenarios surveyed, the respondents prefer the version of the code without the atom of confusion, and in no case the \emph{neutral} ratio was higher than the option for the version without atoms of confusion.An example entry is contained in Appendix~\ref{}. 
For a full listing of the code snippets, visit the paper repository.

\subsubsection*{Participants views of the atoms} Interesting, only for the \emph{ternary operator} scenario the participants prefer the version of the code with the atom of confusion, instead of the more cleaner version without the atom of confusion. Even in this case, some participants who opted for the \emph{left-hand-side} (with the atom of confusion) version still believed that the \emph{right-hand-side} (without the atom of confusion) version was more readable (see Figure~\ref{code:ternary}). 
The following quotes were extracted from three interviews:

@@

\begin{mq}
\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers".}
\end{mq}

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand"}.
\end{mq}

@@

This particular atom of confusion also opens up the possibility for a derivative construct that JavaScript allows which can be rather confusing, and that is the nested ternaries construct, in which the right-hand side of the a ternary operator can be another ternary construct. While nested ternaries remove the number of lines that would be necessary to construct using nested if-then-else statements, they can become quite taxing, cognitively speaking, to understand. Therefore, nested ternaries are a choice of atoms to be analysed in future work.

\rb{Poder\'{i}amos quantificar a existencia de emph{nesting} com code queries?}

The \emph{Pre-Increment} atom of confusion also caused such divide in one of the interviewees, as they regarded the \emph{\rhs version} (without atom of confusion) as simpler to understand, but would still opt to write code that contained the atom. In contrast to such opinion, one of the participants found the version that contained the atom of confusion more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.

When analyzing the \emph{Logic as Control Flow} atom of confusion, one of the interviewees gave an example of his own experience on why one should avoid writing code with such atom:

@@

Two of the atoms had unanimous preference for the versions that did not contain the atoms: \emph{Comma Operator} and \emph{Omitted Curly Braces and Indentation}. Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants. Two remarks about the \emph{Comma Operator} atom of confusion are listed below:

\begin{mq}
\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked"}
\end{mq}

\begin{mq}
\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang"}
\end{mq}

As for the \emph{Omitted Curly Braces and Indentation} atom of confusion, which, as we shall see in Section~\ref{sec:msr-results}, is frequent in JS code bases, one of the interviewees mentioned that, although they understand why one would opt to not use braces for simple if-then-else statements, he still advised against it, on grounds that:

\begin{mq}
\emph{``[I prefer the \rhs version of the code \ldots] If I want to see well-written, easily understandable code, then I also have to do my job. Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code"}
\end{mq}

This view is central to our research on why such small, yet tricky, constructs should be avoided whenever possible, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.

\rb{que conclus\~{a}o extra\'{i}mos com essa parte do estudo?}

\subsection{Common Remarks From the Open-ended Questions} The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.

One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code that, particularly when involving  deep prototype chains. This is a core feature of the language, and most high-level tools and frameworks abstract it away. Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.

When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate. One of the respondents regarded nested callbacks difficult even to write:

\begin{mq}
\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them."}
\end{mq}

@@

A total of 40\% (n = 6) practitioners regarded the JavaScript language as quick and/or easy to learn, and 26.7\%(n = 4) mentioned that the abundance of frameworks, coupled with the abstractions they provide, make it so that one can develop quickly in the language. However, as mentioned by one of the interviewees, ``\emph{Developers who have started programming in JavaScript might miss out on important lower-level aspects, such as garbage collection, memory management, and processor management.}''.

Moreover, such abstractions can make it so that one starts to program using the language without actually understanding some of its basic aspects. For instance, 26.7\%(n = 4) said that the language is not typed, 13.3\%(n = 2) believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements. For the multi-purpose language that is one of the most, if not the most used one by practitioners
~\cite{}, a lack of understanding of the languages core features may be a problem worth investigating in future studies.

@@

\begin{table}[!htb]
    \centering
    \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results2}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Property access         & 10  & 60  & 30  \\
         Object spread           & 50 & 50     & 0    \\
         Array spread            & 60  & 40    & 0     \\
         Arrow function          & 40 & 40     & 20  \\
         Array destructuring     & 40 & 40  & 20  \\
         Object destructuring    & 40 & 40     & 20    \\
         Type conversion         & 30  & 60    & 10     \\
         Change Literal encoding & 40 & 50  & 10  \\ \midrule
         \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
         \bottomrule
    \end{tabular}
\end{table}

\subsection{Results of the third study: Mining open source repositories}
\label{sec:msr-results}

>>>>

%% \section{Results and Discussion}
%% \label{sec:results}

%% In this section we present the main findings of our research. We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}). After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.

\section{Results of the Survey}
\label{sec:survey-resuts}

Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
and \emph{cognitive effort} (time necessary to provide a correct answer). We receive full answers to our survey
from 140 participants. Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects,
according to their education level and years of programming experience, respectively.

\begin{figure}[htb]
      \centering
      \includegraphics[width=\columnwidth]{images/dem-education-1.pdf}
      \caption{Participants' Education Level}\label{fig:degree}
  \end{figure}

  \begin{figure}[htb]
      \centering
      \includegraphics[width=\columnwidth]{images/dem-experience-1.pdf}
      \caption{Participants' Experience} \label{fig:xp}
  \end{figure}

Figure~\ref{fig:xp} shows that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming. No subject reported to have never attended university. Our audience comprises skilful developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}). Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%). Other 37\% of the respondents have between one and four years of experience. In this way, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.

\subsection{Misunderstanding Analysis}

\subsubsection*{Exploratory Data Analysis}

Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. We collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). Our final dataset consists of 70 complete Latin Squares.

Regarding \emph{correctness}, Table~\ref{results_correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey. Accordingly, the version of seven code snippets without atoms of confusion present at least a 15\% improvement in answer correctness. In particular, \emph{Comma Operator} atom presents the highest impact on misunderstanding. 
Curiously, frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.

Moreover, as the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when observing the cleaned version of the code snippets. Also, the sample of answers where there was no atoms had almost no dispersion, which is a sign that the non-confusing code is easier to evaluate correctly.

@@

\begin{table*}[htbp]
\caption{Total number of correct answers.}
\label{tab:difference-correctness}
\centering{
\begin{tabular}{lccc|c} \toprule
 % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
  Atom & Confuse Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
  Comma Operator &  28 &  65 & +132 & 19.02 \\ 
  {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
  Post Increment &  48 &  64 & +33 & 4.83 \\ 
  Ommitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
  Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
  Implict Predicate &  58 &  68 &  +17 & 6.94 \\ 
  Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
  Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
  Pre Increment &  50 &  53 &  +6 & 1.24 \\ 
  Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
\end{tabular}
}
\end{table*}

 %% Comma Operator          & 40 & 93  & +132 \\
 %% Post Increment          & 69 & 91  & + 31  \\
 %% Omitted Curly Braces    & 67 & 83  & +23 \\
 %% Assignment as Value     & 80 & 97  & +21 \\
 %% Implicit Predicate      & 83 & 97  & +16 \\
 %% Logic as Control Flow   & 59 & 68  & +15 \\
 %% Ternary Operator        & 86 & 94  & +9  \\
 %% Pre-Increment           & 71 & 76  & +7  \\
 %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule

\begin{figure}[htb!]
\noindent
 \centering
 \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
 \caption{Number of wrong answers of each subject}
 \label{fig:boxplotcorrectness}
 \end{figure}

\rb{verificar a consistencia nas questoes de pesquisa mais especificas}

\subsubsection*{Hypothesis Testing and Effect Size}

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference
in the frequency of correct and incorrect answers---due to the versions
of the code snippets (confuse and clean code). The result shows that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$). We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
\emph{education} or participant \emph{experience} impact on correctness. Although the results suggest that experience impact correctness, the education level does not. We then replicate the \emph{Pearson's Chi-squared test} for each experience group (see Figure~\ref{fig:xp}) and found that the statistic difference is less significant for those developers with more than ten years of experience.
Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
method (see Table~\ref{tab:difference-correctness}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre Increment. Nonetheless, for the remaining candidates, the effect size is considerable. For instance, the Comma Operators leads to an \emph{Odds Ratio} of
\num{19.02}, meaning that the participants are \num{19.02} times
more likely to correctly interpret the clean version of a code in comparison with the corresponding, confuse version of the code snippet.

%% Regarding the first question we 
%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code. We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).

\subsection{Cognitive Workload Analysis}

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants give a correct answer about the expected outcomes of a code snippet, considering both confuse and clean versions. Eight atom candidates require an increasing cognitive workload. For these eight atom candidates, developers take at least 5.90\% less time on average to find a correct answer---when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, take less time to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, there was an increase
of 29.06\% and 38.19\% in the average time taken to give a correct answwer---wherein the clean code versions.

\begin{table*}[htbp]
\caption{Time in seconds to submit a {\color{red}correct} answer}
\label{tab:difference-time-taken}
\centering{
\begin{tabular}{lccc|cc} \toprule
Atom & Confuse Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
{\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
Ommitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
Implict Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
Pre Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
         \bottomrule
    \end{tabular}}
\end{table*}

\subsubsection*{Hypotheis Testing}
We use the non-parametric \emph{Mann-Whitney test} to
investigates the null hypothesis that developers 
spend the same ammount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confuse or the clean version
of the code. Table~\ref{tab:difference-time-taken}
shows the results, which suggest that we
should refute the null hypothes for
five atom candidates (Comma Operator, Logic
as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre Increment). For the first three,
the analysis suggest that the participants
need more time to predict the outcome of
the code snippet in the confuse code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre Increment, the participants take less time
to predict the outcome of the code snippets in
the confuse version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre Increment.

\rb{Acho que dever\'{i}mos justificar o uso do Cliff's Delta aqui e o Odds Ratio na
  se\c c\~{a}o anterior.}

\subsection{Discussion}

\rb{Acho que podemos incluir uma discuss\~{a}o ao
  t\'{e}ermino de cada uma das se\c c\~{o}es com
  resultados. Ou separamos em uma \'{u}nica
se\c c\~{a}o de discuss\~{a}o?}

%%  Comma Operator        & 60 & 21  & -65  \\
%% Logic as Control Flow & 85 & 49  & -42  \\
%% Implicit Predicate    & 33 & 24  & -27  \\
%% Omitted Curly Braces  & 43 & 31  & -27  \\
%% Assignment as Value   & 53 & 49  & -7   \\
%% Ternary Operator      & 42 & 42  & 0    \\
%% Post Increment        & 27 & 27  & 0    \\
%% Arithmetic as Logic   & 29 & 36  & +24  \\
%% Pre-Increment         & 34 & 49  & +44  \\

\
% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

\section{Results of the Interview Study}
\label{sec:interview-results}

\rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
  gerando uma complexidade desnecess\'{a}ria. vou tentar
  remover isso}.

In this section we present the results of the 
interviews with the practitioners who agreed to participate
in the study. A total of 15 practitioners
took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the {\color{red}nine} atom candidates---by
(a) presenting them pairs of code snippets whose behavior were identical
and then (b) asking them to compare the code snippets readability.

%% In the second round of interviews, ten of the fifteen developers
%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase. The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.

Supporting the results of the survey, Tables~\ref{tab:interview-results1}
and~\ref{tab:interview-results2} show that for twelve out of the
seventeen scenarios surveyed, the respondents prefer the version of the
code without the atom of confusion, and in no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository.

\subsubsection*{Participants views of the atoms} Interesting,
only for the \emph{ternary operator} scenario the participants
prefer the version of the code with the atom of confusion,
instead of the more cleaner version without the atom of confusion.
Even in this case, some participants who opted for
the \emph{left-hand-side} (with the atom of confusion) version
still believed that the \emph{right-hand-side} (without the atom of
confusion) version was more readable (see Figure~\ref{code:ternary}). 
The following quotes were extracted from the transcripts with
three inteviwes:

@@

\begin{mq}
\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers''}
\end{mq}

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand''}.
\end{mq}

@@

The atom candidate Ternary Operator also opens up
the possibility for a derivative construct that JavaScript
allows which can be rather confusing, and that is the nested
ternaries construct, in which the right-hand side of the a
ternary operator can be another ternary construct.
While nested ternaries remove the number of lines that
would be necessary to construct using nested if-then-else statements,
they can become quite taxing, cognitively speaking, to understand.
Therefore, nested ternaries are a choice of atoms to be analysed in future work.

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, as they
regarded the \emph{\rhs version} (without atom of confusion)
as simpler to understand, but would still opt to write
code that contained the atom. In contrast to such opinion,
one of the participants found the version that contained the atom
of confusion more elegant, but recognized it was less readable,
and was willing to sacrifice elegance for readability.

\rb{pela discuss\~{a}o anterior, parece-me que alguns
  desenvolvedores reconhecem que esses candidatos
  a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
  mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
  Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
on why one should avoid writing code using this construct:

@@

Two of the atom candidates had unanimous preference for the
versions that did not contain the atoms: Comma Operator and
Omitted Curly Braces. Regarding the first one, we could
often notice during the interviews that
the \emph{\lhs} (with the atom of confusion) caused significant
confusion among the participants. Two remarks about
the \emph{Comma Operator} atom of confusion are listed below:

\begin{mq}
\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
\end{mq}

\begin{mq}
\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}
\end{mq}

As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, he still advised against it, on grounds that:

\begin{mq}
\emph{``[I prefer the \rhs version of the code \ldots] If I want to see well-written, easily understandable code, then I also have to do my job. Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}
\end{mq}

This is an interesting perspective, because such small, yet tricky, constructs should be avoided, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.

% \subsection{Common Remarks From the Open-ended Questions}

The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.

One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving  deep prototype chains. This is a core feature of the language, and most high-level tools and frameworks abstract it away. Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.

\rb{seria importante incluir uma cita\c c\~{a}o para a \'{u}ltima afirma\c c\~{a}o acima.}

When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate. One of the respondents stated:

\begin{mq}
\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them.''}
\end{mq}

@@

Six practitioners regarded the JavaScript language as easy to learn, while four
participants mentioned that the abundance of frameworks,
coupled with the abstractions they provide, make it so that one
can develop quickly in the language. However, as mentioned by one
of the interviewees, ``\emph{Developers who have started
programming in JavaScript might miss out on important lower-level
aspects, such as garbage collection, memory management,
and process management.}''.

\rb{esse pr\'{o}ximo par\'{a}grafo eh interessante,
  se refor\c carmos que os desenvolvedores tem
  experi\^{e}ncia com a linguagem. caso contr\'{a}rio,
  parece que nao fizemos uma boa selecao de participantes.
  algo para discutirmos.}

Moreover, such abstractions might lead a developer to start to program
using the language without actually understanding some of its basic aspects.
For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements. For the multi-purpose language that is one of the most, if not the most used one by practitioners
~\cite{}, a lack of understanding of the languages core
features may be a problem worth investigating in future studies.

@@

%% \begin{table}[!htb]
%%     \centering
%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
%%     \label{tab:interview-results2}
%%     \begin{tabular}{lrrr}\toprule
%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
%%       \cmidrule(lr){2-4}
%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
%%          Property access         & 10  & 60  & 30  \\
%%          Object spread           & 50 & 50     & 0    \\
%%          Array spread            & 60  & 40    & 0     \\
%%          Arrow function          & 40 & 40     & 20  \\
%%          Array destructuring     & 40 & 40  & 20  \\
%%          Object destructuring    & 40 & 40     & 20    \\
%%          Type conversion         & 30  & 60    & 10     \\
%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
%%          \bottomrule
%%     \end{tabular}
%% \end{table}

\section{Results of the MSR Effort}
\label{sec:msr-results}
